MachLensFinder - I: An automated Strong Lensing Identification Pipeline

### 2016-12-19 ms.tex
Only two or three paragraphs to go in the intro, and citations

\begin{document} 
\shorttitle{LensFinder}
\shortauthors{Avestruz, Li, Lightman}
\submitted{The Astrophysical Journal} 
\slugcomment{The Astrophysical Journal, submitted}

\title{MachLLensFinder - I: A Machine Learning Strong Lensing
  Identification Pipeline}

\author{Camille Avestruz\altaffilmark{1-3}\thanks{E-mail:
    avestruz@uchicago.edu}, Nan Li\altaffilmark{2-4}, and Matthew Lightman\altaffilmark{5} }

\affil{
$^1${Enrico Fermi Institute, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^2${Kavli Institute for Cosmological Physics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^3${Department of Astronomy \& Astrophysics, The University of Chicago, Chicago, IL 60637 U.S.A.};\\
$^4${High Energy Physics Division, Argonne National Laboratory,
Lemont, IL 60439};\\
$^5${Digital Intelligence, JPMorgan Chase, Chicago, IL 60603 U.S.A.};\\
  \href{mailto:avestruz@uchicago.edu}{avestruz@uchicago.edu}\\
}

@@

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present \change{LensFinderML}, a strong
lensing identification pipeline that will be publicly released as open
source software.  In this first application, we employ a fast feature
extraction method, Histogram of Oriented Gradients (HOG), to capture
edge features that are characteristic of strong gravitational arcs.
We use logistic regression to train a supervised classifier model on
the HOG of \change{HST- and LSST}-like images.  Our best performing
models reach a completeness rate of \change{XX\%} and purity rate of
\change{XX\%} for an HST-like sample.  However, for one orbit of LSST,
our model cannot distinguish between strong lensing systems and
non-strong lensing systems.  In absence of the brightest cluster
galaxy in LSST-like data, our trained model reaches a completeness of
\change{XX\%} and purity of \change{XX\%}.  Our results show one the
most efficient and effective methods for automated strong lensing
identification in space-based surveys.  The results also emphasize the
need for further development and testing of additional tools to
exploit[wc?]  ground-based surveys.
\end{abstract}
%-------------------------------------------------%

@@

% Gravitational lensing as a geometric (?) test of cosmology.  Also
% probes underlying matter distribution in galaxy clusters, sensitive
% to things like recent accretion and environment.
Gravitational lensing deflections due to gravity... \citep[see][for a
  review]{kneibandnatarajan_11}.  Strong lensing visible ... giant
arcs, multiple images, and arclets.  Galaxy clusters as strong lens,
deep gravitational potential well, can see by eye.  Strong lensing
signatures probe underlying dark matter distribution of galaxy
clusters, magnify the background galaxy population allowing us to
probe formation at epochs further back in time, and geometric test of
cosmology.

% Arc identification can be done visually (e.g. that zooniverse
% thing).  But, in the advent of optical surveys, DES and LSST, we
% will be limited by our ability to automatically process the large
% amounts of data.  Wide field surveys - SPT in millimeter, Herschel
% in submillimeter; high resolution spectroscopic imaging with ALMA;
% optical surveys DES, HSC, LSST, and Euclid (Oguri \& Marshall 2010).
The geometric test of cosmology requires a comparison of predicted arc
abundances and observed abundances.  In the advent of large surveys,
such as DES and LSST, we will be limited by our ability to
automatically process the large amounts of data and select likely
strong lensing clusters for follow up spectroscopic measurements to
positively identify background galaxy images.  Image identification
has had an onslaught of new works (wc).

% Serentipitous discoveries confirmed the theory (cite Einstein)
% Cite first discoveries of different shapes of strong lensing images
% While some serendipitous systems found recently (gladders+),
% Describe the uptick in amount of data
% End with need to scale with the volume of data
Early strong lensing images were serendipitous discoveries
(e.g. multiple quasars \citep{walsh_etal79}, arcs in clusters
\citep{lyndsandpetrosian_86}, ...).  As the number of images with
potential lensing systems increased, infrastructure emerged for both
automated and visual inspection.

% Image identification works: visual, ``robot'' identification -
% ArcFinder, RingFinder..., and citizen science (space warps).
% Many of these rely on some explicit aspect of the image's morphology (give example), filtering, ... ....
% Specifics of the methodology how the instrument affects the strong lensing signal, and the objects of interest.

Automated lensed image identification has fallen under the term
``robot'' identification, where images are fed into some kind of image
processing pipeline that enhances and extracts characteristic features
as parameters that are fed into some kind of cutoff or pattern
recognition piece.  ArcFinder
... .... \citep{seidelandbartelmann_07}. For multiply imaged quasars,
Ringfinder ... ... ...\citep{gavazzi_etal14}.  Each pipeline
.... ... and relies on the image morphology where ... ...  Note, that
the method best applied to ... ... and the objects of interest.  The
human eye is one of the best discriminators of.... SpaceWarps is an
example of citizen science... visual identification
\citep{marshall_etal16,more_etal16}.  However, scalability, training
....

% Arc finding must optimize for completeness and purity.  Completeness
% (no false negatives), purity (minimal false positives).  Purity can
% be double checked in a smaller subsample of the data visually.  And,
% spectroscopic follow up is expensive.  Also, scalability of the
% strong lens finding method matters - example with the citizen
% science.
In general, lensed image finding must optimize for completeness and
purity.  In the application of arc abundance comparisons with
cosmological predictions, completeness is important; we must be able
to understand the reasons and frequency of false negatives.  Since
spectroscopic confirmation of background lensed images is expensive,
purity is key; we must minimize false positives that would waste
follow up telescope time.

% Machine learning to automate.  Previous works (StrongML),
% More+... Here, we focus on the improvement from preprocessing the
% images and describe the pipeline, which is publicly available.
% Depending on the image type, different preprocessing is likely
% necessary (rings, multiply lensed quasars, etc.) Our goal is to
% create an open source pipeline that is general enough for the user
% to select or add appropriate image processing techniques to augment
% the features of the lensed image, train and test data with different
% machine learning techniques, and to select the sequence of methods
% that maximizes completeness with optimal purity.

% No need to explicitly extract morphological parameters to a pattern
% finding, simply transform the image according to the appropriate
% image preprocessing function, and directly feed this into the
% machine learning ...

% This is the first in a series of LensFinder papers where we present
% our initial pipeline results with a test... of .... Image
% preprocessing.... same technology to validate watermarks

Our paper is organized as follows.  In Section~\ref{sec:methods} we
briefly describe ...  We present our results in
Section~\ref{sec:results}, and our summary and discussions in
Section~\ref{sec:conclusions}.

>>>>

\begin{document} 
\shorttitle{MachLensFinder}
\shortauthors{Avestruz, Li, Lightman}
\submitted{The Astrophysical Journal} 
\slugcomment{The Astrophysical Journal, submitted}

\title{MachLensFinder - I: An automated Strong Lensing
  Identification Pipeline}

\author{Camille Avestruz\altaffilmark{1-3,$\dagger$}\thanks{E-mail:
    avestruz@uchicago.edu}, Nan Li\altaffilmark{2-4}, and Matthew Lightman\altaffilmark{5} }

\affil{
$^1${Enrico Fermi Institute, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^2${Kavli Institute for Cosmological Physics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^3${Department of Astronomy \& Astrophysics, The University of Chicago, Chicago, IL 60637 U.S.A.};\\
$^4${High Energy Physics Division, Argonne National Laboratory,
Lemont, IL 60439};\\
$^5${Digital Intelligence, JPMorgan Chase, Chicago, IL 60603 U.S.A.};\\
$^\dagger${Provost's Postdoctoral Scholar at the University of Chicago}\\
  \href{mailto:avestruz@uchicago.edu}{avestruz@uchicago.edu}\\
}

@@

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present {\em MachLensFinder}, a
strong lensing identification pipeline that will be publicly released
as open source software.  In this first application, we employ a fast
feature extraction method, Histogram of Oriented Gradients (HOG), to
capture edge patterns that are characteristic of strong gravitational
arcs in galaxy-galaxy lensing.  We use logistic regression to train a
supervised classifier model on the HOG of \change{HST- and LSST}-like
images.  Our best performing models reach a completeness rate of
\change{XX\%} and purity rate of \change{XX\%} for an HST-like sample.
However, for one orbit of LSST, our model cannot distinguish between
strong lensing systems and non-strong lensing systems.  In absence of
the brightest cluster galaxy in LSST-like data, our trained model
reaches a completeness of \change{XX\%} and purity of \change{XX\%}.
Our results show one the most efficient and effective methods for
automated strong lensing identification in space-based surveys.  The
results also emphasize the need for further development and testing of
additional tools to maximize use of ground-based surveys.
\end{abstract}
%-------------------------------------------------%

@@

% Gravitational lensing as a geometric (?) test of cosmology.  Also
% probes underlying matter distribution in galaxy clusters, sensitive
% to things like recent accretion and environment.
Gravitational lensing occurs when intermediate fluctuations in the
matter density field deflect light from background sources
\citep[see][for a review]{kneibandnatarajan_11}.  In particular, the
deep gravitational potential wells of galaxy clusters provide a strong
lens, with lensed images that can be seen by eye.  Strong
gravitational lensing can manifest as visible giant arcs magnifying
high redshift galaxies \citep{lyndsandpetrosian_86,gladders_??},
multiply imaged quasars \citep{walsh_etal79}, and arclets.  Lensing
signatures probe the underlying dark matter distribution of galaxy
clusters \citep{warrenanddye_03}, probe high redshift galaxy formation
\citep{allam_etal07}, and provide a geometric test of cosmology.

% Arc identification can be done visually (e.g. that zooniverse
% thing).  But, in the advent of optical surveys, DES and LSST, we
% will be limited by our ability to automatically process the large
% amounts of data.  Wide field surveys - SPT in millimeter, Herschel
% in submillimeter; high resolution spectroscopic imaging with ALMA;
% optical surveys DES, HSC, LSST, and Euclid (Oguri \& Marshall 2010).
The geometric test of cosmology requires a comparison of predicted arc
abundances and observed abundances
\citep{kochanek_96,chae_02,linder_04}.  Visual arc identification has
been effective through the use of citizen science platforms
(e.g. \citet{marshall_etal16,more_etal16}).  However, in the advent of
large surveys, such as the Dark Energy Survey (DES) and the Large
Synoptic Sky Telescope (LSST), we will be limited by our ability to
automatically process large amounts of data and select likely strong
lensing clusters for follow up visual and spectroscopic confirmation.
The strong lensing sector of cluster-based cosmology requires an
exploration of techniques in image processing and classification.

% Image identification works: visual, ``robot'' identification -
% ArcFinder, RingFinder..., and citizen science (space warps).  Many
% of these rely on some explicit aspect of the image's morphology
% (give example), filtering, ... ....  Specifics of the methodology
% how the instrument affects the strong lensing signal, and the
% objects of interest.
Over the last decade, infrastructure has emerged for both large scale
visual and automated image classification. By nature, the human eye is
one of the best discriminators for image classification.  {\em
  SpaceWarps} is an example of citizen science based image
classification of strong lensing systems in Canada France Hawaii
Telescope Science (CFHTLS) telescope observations
\citep{marshall_etal16,more_etal16}.  Visual classification has been
quite successful for a dataset like CFHTLS, classifying \change{XX in
  XX time}.  However, future datasets like Euclid and LSST will be
\change{XX} times this size, challenging the scalability of a citizen
science approach.  The next step is to find automated methods with
performance comparable to or better than humans. {\em SpaceWarps} is a
part of the Zooniverse Project (CITE???), which also includes {\em
  Galaxy Zoo}, the citizen science based image classification of
galaxy types (CITE???).  Galaxy classification is a successful example
of in which machine learning algorithms successfully train models to
classify images with comparable performance to humans (CITE???).

% Describe general robot identification schemes that have been
% implemented
General automated, or ``robot'', identification refers to a classifier
where images are fed into an image processing pipeline that enhances
and extracts characteristic features as parameters on which a pattern
recognition piece is applied [wc???].  Examples of this include
ArcFinder ... .... \citep{seidelandbartelmann_07}. For multiply imaged
quasars, Ringfinder ... ... ...\citep{gavazzi_etal14}.  Each pipeline
.... ... and relies on the image morphology where ... ...  Note, that
the method best applied to ... ... and the objects of interest.

% Describe how machine learning is a step up
With ``machine learning'' algorithms, we can train a model to separate
a dataset according to some desired classification scheme.  However,
we still need to reduce the dimensionality of the problem by removing
information in the images that do not pertain to the classifier.
Examples of automated classification will...

% Previous works and our goal to make a universal pipeline 
Machine learning to automate.  Previous works (StrongML),
More+... Here, we focus on the improvement from preprocessing the
images and describe the pipeline, which is publicly available.
Depending on the image type, different preprocessing is likely
necessary (rings, multiply lensed quasars, etc.) Our goal is to create
an open source pipeline that is general enough for the user to select
or add appropriate image processing techniques to augment the features
of the lensed image, train and test data with different machine
learning techniques, and to select the sequence of methods that
maximizes completeness with optimal purity.

% This is the first in MachLensFinder series with initial pipeline
% results
This is the first in a series of {\em MachLensFinder} papers.  We
present our image processing and machine learning classifier pipeline.
Our image processing procedure parameterizes edges in our image in a
1-d histogram, enabling fast feature extraction that is similar to
techniques used to identify humans in security software systems.  We
train a model with logistic regression to identify features
corresponding to strong lensing systems.

We show results of the pipeline on mock galaxy-galaxy lens systems
observed by HST and LSST as respective examples of classifier
performance on optical space- and ground-based observations.
\citet{miraldaescudeandlehar_92} first suggested that massive
ellipticals would be likely frequent strong lensing sources in optical
surveys.  We train and test our pipeline on subsamples from 10,000
mock observed brightest cluster galaxy (BCG) strong lensing systems
and 10,000 BCG non-strong lensing systems.

% Arc finding must optimize for completeness and purity.  
In general, lensed image finding must optimize for completeness and
purity.  In the application of arc abundance comparisons with
cosmological predictions, completeness is important; we must be able
to understand the reasons and frequency of false negatives.  Since
spectroscopic confirmation of background lensed images is expensive,
purity is key; we must minimize false positives that would waste
follow up telescope time.  We discuss the performance of our image
processing and classification pipeline with respect to the
completeness and purity rates of our trained models and illustrate the
scalability of training our models [wc???].

Our paper is organized as follows. In Section~\ref{sec:methods} we
briefly describe the methods to generate the mock HST and LSST data
and our overall image processing and classification pipeline. We
present our results in Section~\ref{sec:results}, and our summary and
discussions in Section~\ref{sec:conclusions}.
